Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-02 document/ms.tex
clearer description of importance sampling approximation

The quantity that we need to compute is the \emph{marginalized likelihood} of
the data \obs{\catalog}\ given a particular setting of the rate function
parameters \ratepars.
This is achieved by marginalizing over the true periods and radii
\begin{eqnarray}
p(\obs{\catalog}\,|\,\ratepars) &=&
\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \\
&=&
\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog) \dd\catalog
\end{eqnarray}
where the second line is true because the rate function is defined on the true
parameters.
This integral can then be written in terms of the sampling under interim
priors
\begin{eqnarray}
p(\obs{\catalog}\,|\,\ratepars)
&=&
\int p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)\,
\frac{p(\catalog\,|\,\obs{\catalog},\,\interim)}
     {p(\catalog\,|\,\obs{\catalog},\,\interim)}
\dd\catalog
\\
&=&
p(\obs{\catalog}\,|\,\interim)\,
\int \frac{p(\catalog\,|\,\ratepars)}{p(\catalog\,|\,\interim)}\,
p(\catalog\,|\,\obs{\catalog},\,\interim)
\dd\catalog \quad.
\end{eqnarray}
We can approximate this integral using the samples $\catalog^{(n)}$ defined
above and find the marginalized likelihood up to an unimportant normalization
constant
\begin{eqnarray}\eqlabel{marg-like}
p(\obs{\catalog}\,|\,\ratepars) &\appropto&
\sum_{n=1}^N
\frac{p(\catalog^{(n)}\,|\,\ratepars)}{p(\catalog^{(n)}\,|\,\interim)}
\end{eqnarray}
where the numerator inside the sum is given by \eq{true-like} and the
denominator is the interim prior evaluated at the samples.

In practice, the variance of this approximation depends on the choice of
interim prior and, in particular, it is useful to include the completeness
function.
\dfmtodo{BLAH BLAH BLAH.}

>>>>

Now that we have a representation of the catalog, we want to infer constraints
on the occurrence rate of exoplanets conditioned on all the data.
To perform this inference, we need to compute the  \emph{marginalized
likelihood}
\begin{eqnarray}\eqlabel{marg-likelihood}
p(\obs{\catalog}\,|\,\ratepars) &=&
\int p(\catalog,\,\obs{\catalog}\,|\,\ratepars) \dd\catalog \quad.
\end{eqnarray}
The marginalized likelihood measures the probability of all the \kepler\ light
curves given a particular setting of the occurrence rate parameters.
If we can efficiently compute this then we can optimize to find the
maximum likelihood occurrence rate or we can apply priors $p(\ratepars)$ and
sample to find posterior constraints on the occurrence rate.
Note that any results obtained through \eq{marg-likelihood} will
\emph{correctly account for all the measurement uncertainties}.

The integral in \eq{marg-likelihood} cannot be computed analytically and, in
fact, it will generally be extremely computationally expensive to evaluate
numerically (\dfmtodo{CITE}).
In our case, however, we only need to evaluate this integral up to a constant
and there is a simple approximation---using the sampling representation of the
catalog---that can be sufficiently accurate for our purposes.
To derive this approximation, first recall that the joint distribution inside
the integral in \eq{marg-likelihood} factorizes as
\begin{eqnarray}
p(\catalog,\,\obs{\catalog}\,|\,\ratepars) &=&
p(\catalog\,|\,\ratepars)\,p(\obs{\catalog}\,|\,\catalog)
\end{eqnarray}
because the occurrence rate generates the \emph{true} distribution of periods
and radii.
Then, we can multiply the integrand by
\begin{eqnarray}
\frac{p(\catalog\,|\,\obs{\catalog},\,\interim)}
     {p(\catalog\,|\,\obs{\catalog},\,\interim)}
\end{eqnarray}
and simplify to find
\begin{eqnarray}
\frac{p(\obs{\catalog}\,|\,\ratepars)}
     {p(\obs{\catalog}\,|\,\interim)}
&=&
\int \frac{p(\catalog\,|\,\ratepars)}{p(\catalog\,|\,\interim)}\,
p(\catalog\,|\,\obs{\catalog},\,\interim)
\dd\catalog \quad.
\end{eqnarray}
Using the samples over catalogs $\catalog^{(n)}$ as defined above, we can
approximate this integral---up to an irrelevant normalization constant---as
\begin{eqnarray}\eqlabel{approx-marg-likelihood}
p(\obs{\catalog}\,|\,\ratepars) &\appropto&
\sum_{n=1}^N
\frac{p(\catalog^{(n)}\,|\,\ratepars)}{p(\catalog^{(n)}\,|\,\interim)}
\end{eqnarray}
where the numerator inside the sum is given by \eq{true-like} and the
denominator is the interim prior evaluated at the coordinates of the samples.

This is a generalization of the approximation presented previously in the
exoplanet literature (\citealt{hogge}) but in this case, the likelihood
function \eq{true-like} doesn't factorize over entries in the catalog.
As a result, the variance of \eq{approx-marg-likelihood} is more sensitive to
the choice of interim prior.
In practice, we have found that including the completeness function
$\completeness(\entry)$ in $p(\entry\,|\,\interim)$ greatly improves
performance.
This term can be incorporated into existing samplings---generated using less
informative priors---using rejection sampling.

\section{Parameterization of the occurrence rate}
